MS ANXIETY

### 2025-01-23 scripts/jama_scripts/MSAnxiety_manuscript_JAMA_revision_20250123_FINAL.Rmd
Add files via upload

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
library(MatchIt)
library(ggpubr)
library(tidygam)
library(psych)
#library(tidymv)
library(aod)
library(viridis)  # Ensure you have the viridis package
#functions
source("~/Library/CloudStorage/Box-Box/BBL/ms_generic_scripts/msdepression_functions.R")
```

## MS anxiety analyses

## R Markdown

The first chunk here is reading in appropriate data frames and preprocessing. Two big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller. It does however highlight the

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

Inclusion criteria below:

Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
  1) Scanned under MS protocol (42K)
  2) MS Diagnosis (17K)
  3) Seen by MS doctor (16K)

Inclusion criteria for Anxious Group
  1) Have an ICD10 F4* code (in our sample, people have F41.1; 41.3; 41.8; 41.9)
        (may use this later)- https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

Inclusion criteria for Healthy Group
  1) No F* diagnoses
  2) At least 1 PHQ2 or PHQ9 with a documented 0
  3) Free of psychiatric medications per NAMI website (plus editions per Dr. Baller)

Exclusions
  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of anxiety. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

Imaging exclusion
  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.

Groups:
1. Total unique individuals who also have MS - 1026
2. Total unique who had good MIMoSA - 867 (lost 159)
3. Remove people without MS provider, n=814 (lost 53)

#figure brainstorming
1. Flowchart of inclusion/exclusion (825 participants -> 11 lost for poor streamline filtering -> 814 go into phenotyping) -> Elena
   A. Anxiety diagnosis
    I. Yes
       - Anxiety Med
          a. Yes -> MS+severeA (24 individuals)
          b. No  -> MS+mildA (87 indiv)
     II. No
        - Anxiety Med
          a. Yes -> MS+mildA (162)
          b. No
            - Has PHQ2 or 9 and 0 
              i. Yes-> MS+noA (99 individuals)
              ii. No -> Unclassified (442 individuals)

Final groups (MS+noA = 99, MS+mildA = 249, MS+severeA = 24)

2. Demographics (age, sex, race, +/- PHQ2, PHQ9, has anxiety dx, on anxiety meds, both, on dmt) -> elena to make all the csvs into tables
3. PROMIS scores (each type of measure is a line, y axis is score, x axis is anxiety dose)
4. Some sort of schematic of pipeline (Maggie) - 
  a. amygdala - mpfc blobs
  b. uncinate connecting the two
  c. white matter lesion dropped in uncinates
  d. streamline filtering
  e. mean of both side
5. Mean Anxiety dose either violin or bar

6. Supplement: 
 - Promis tables
#Rstduio Version: 2024.04.1+748 (2024.04.1+748)
## Make final data frame for analysis
### read in emr data
```{r read_in_df}
homedir <- "~/Library/CloudStorage/Box-Box/BBL/msanxiety/"

@@

#save 
simple_demo_diagnosis_df <- print(simple_demo_diagnosis, as.data.frame = TRUE)
write.csv(simple_demo_diagnosis_df, file = paste0(homedir, "/results/simple_demo_diagnosis_df.csv"), row.names = TRUE)
```
### MS med summaries by group
```{r demographics}

@@

#####Promis x anxiety dose, group by mental vs physical health
```{r plot_Promis_x_anxiety}
#set mood versus physical domains
mood_domains <- c("Mental_Health_and_Mood",
                "Social_Activities_Satisfaction",
                "Carrying_Out_Social_Activities",
                "Emotional_Problems", 
                "Quality_of_Life",
                "PostOp.PROMIS.Mental.Score")

@@

```
####Test promis*anxiety dose  stats
#####NEW JAMA ANALYSIS: Testing for Normality - all normally distributed
```{r normality_testing}
###

@@

### Lesion volumes 
#### Diagnosis - sick > healthy in vol of lesions
```{r lesion_volumes_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, volume_of_mimosa_lesions) %>%
  ungroup() #n=123

@@

#gam lesion volumes; DiagnosisMS+Anxiety   8215.7     2702.7   3.040  0.00292 ** 
volume_of_mimosa_lesions_gam <- gam(volume_of_mimosa_lesions ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(volume_of_mimosa_lesions_gam)

 #print summary stats; (T = 3.04, P = 0.003, Cohen's f2 =  0.342, Effect size = medium)
 gam_summary_stats(volume_of_mimosa_lesions_gam)
```

@@

 #bilat anxiety dose + sex + age: anxiety_dose     3866       1254   3.084   0.0022 **

  anxiety_dose_volume_of_mimosa_lesions_by_sex_gam <- gam(volume_of_mimosa_lesions ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

   #print summary stats; (T = 3.084, P = 0.0022, Cohen's f2 =  0.12, Effect size = small)
 gam_summary_stats(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

@@

### Whole brain volumes 
#### Diagnosis - sick > healthy in vol of lesions
```{r lesion_volumes_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, total_volume) %>%
  ungroup() #n=123

@@

 #print summary stats
 gam_summary_stats(total_brain_volume_gam)
```

#### Anxiety dose
  - anxiety dose correlates with disease burden (total brain volume )
```{r lesion_volumes_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != -1) #n=372

@@

#### Uncinate by anxiety diagnosis (prop)
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

 #print summary stats
  gam_summary_stats(disease_mean_UF_prop_gam)
```

@@

@@

    #bilat anxiety dose + sex + age: anxiety_dose    0.04989    0.02397   2.082   0.0381 * 
 anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

  #print summary stats
  gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)

@@

#### Sensitivity
###### UF by depGroupVar (vol/prop, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the UF findings are specific to anxiety, so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

 #mean UF, depDiagnosisMS+noDepression   -0.05029    0.03028  -1.661   0.0978 . 
dep_mean_UF_gam_prop <- gam(mean_UF_prop ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_UF_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_UF_gam_prop)

@@

###### Lesion & brainvol by depGroupVar (lesion vol - Y, tbv - N, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the lesion and brain vol findings are nonspecific t so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

 #mean vol mimosa lesions, depDiagnosisMS+noDepression 
#(T = -3.164, P = 0.002, Cohen's f2 =  0.128, Effect size = small)
dep_mean_lesion_vol_gam_prop <- gam(volume_of_mimosa_lesions ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_lesion_vol_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_lesion_vol_gam_prop)

  #mean brain vol mimosa lesions, depDiagnosisMS+noDepression  
 #(T = 0.016, P = 0.988, Cohen's f2 =  0.333, Effect size = medium)
dep_mean_total_brain_vol_gam_prop <- gam(total_volume ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_total_brain_vol_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_total_brain_vol_gam_prop)

@@

###### Fornix by Anxiety Diagnosis and Dose (vol and prop, N.S)
```{r fornixByAnxietyDose}

 ##### healthy vs sick - prop
#Fornix gam Diagnosis; DiagnosisMS+Anxiety    0.03690    0.05405   0.683   0.4961    - NS
anxietyDiagnosis_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)

#define group for dose
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1)

  #gam, Fornix anxiety_dose   0.02303    0.02401   0.959    0.338    - NS
  anxietyGrouping_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_F_by_sex_gam_prop)

 #print summary stats
 gam_summary_stats(anxietyDiagnosis_mean_F_by_sex_gam_prop)
 gam_summary_stats(anxietyGrouping_mean_F_by_sex_gam_prop)
```

###### JAMA ANALYSIS- UF Anxiety diagnosis + lesion vol
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

  #how about in fornix

  corr_LV_to_fornix <- cor.test(uncinate_by_dx_df$volume_of_mimosa_lesions,uncinate_by_dx_df$mean_fornix_volume) #corr 0.69; 
  print(corr_LV_to_fornix) #0.3452654

  lesion_vol_fornix_lm <- lm(uncinate_by_dx_df$volume_of_mimosa_lesions~ uncinate_by_dx_df$mean_fornix_volume)
  summary(lesion_vol_fornix_lm) #T = 4.047

  ## Try scaling mean volume
 # disease_mean_scaled_UF_vol_gam <- gam(mean_UF_vol_scaled_total_vol_lesions ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
  #summary(disease_mean_scaled_UF_vol_gam) #- N.S.

  #print summary stats; (T = -0.556, P = 0.579, Cohen's f2 =  0.007, Effect size = minimal)"
  #gam_summary_stats(disease_mean_scaled_UF_vol_gam)

## modeling this doesn't make sense
#disease_mean_UF_prop_w_LV_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F) + volume_of_mimosa_lesions, data = uncinate_by_dx_df)
#summary(disease_mean_UF_prop_w_LV_gam)

 #print summary stats
 # gam_summary_stats(disease_mean_UF_prop_w_LV_gam)
```
###### JAMA ANALYSIS-  UF Anxiety dose + lesion vol
```{r uncinateByAnxietyDose}

@@

corr_LV_to_UF_anxiety_dose <- cor.test(df_demo_and_fascicles_no_unclass_anxiety_dose$volume_of_mimosa_lesions,df_demo_and_fascicles_no_unclass_anxiety_dose$mean_UF_vol) #
  print(corr_LV_to_UF_anxiety_dose)

```
###### JAMA ANALYSIS- UF Anxiety diagnosis + total brain vol
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

###### JAMA ANALYSIS-  UF Anxiety dose + total brain volume
```{r uncinateByAnxietyDose}

@@

@@

@@

### Plots

##### Overall disease in brain
```{r overall_disease_bruden}
#n = 372

df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified")

prop_injured_by_anxiety_burden_collapsed_vars_plot <-  ggplot(df_disease_burden_unc_and_anx_net_collapsed_variables, aes(x= oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed , y=proportion_volume_lost_per_total_network_size_sum, fill= oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed )) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  theme(axis.text.x = element_text(size=8, angle = 45, vjust = 0.5))
prop_injured_by_anxiety_burden_collapsed_vars_plot
```

###Promis data in UF - groups too small
#### Uncinate by anxiety diagnosis (N.S) 
```{r uncinate_by_dx}

### Results:  
#define group
  uncinate_by_dx_df <- df_people_with_promis_data_scores_added_unique %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup() #n=19

#bilateral UF prop ~ Diagnosis: DiagnosisMS+Anxiety -0.02490    0.11510  -0.216   0.8317 
disease_mean_UF_prop_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_prop_gam)

    #print summary stats
 gam_summary_stats(disease_mean_UF_prop_gam)

#### Uncinate by Anxiety Dose (N.S.)
```{r uncinateByAnxietyDose}

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_people_with_promis_data_scores_added_unique %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) #n=64

  #bilat anxiety dose -anxiety_dose   -0.006261   0.055025  -0.114 0.909793       NS
  anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

    #print summary stats

 gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)

```{r ants}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
library(MatchIt)
library(ggpubr)
library(aod)
library(viridis)  # Ensure you have the viridis package
#functions
source("~/Library/CloudStorage/Box-Box/BBL/msanxiety/scripts/FINAL/ms_functions.R")
```

### read in emr data
```{r read_in_df}
homedir <- "~/Library/CloudStorage/Box-Box/BBL/msanxiety/"

@@

#save 
simple_demo_diagnosis_df <- print(simple_demo_diagnosis, as.data.frame = TRUE)
write.csv(simple_demo_diagnosis_df, file = paste0(homedir, "/results/simple_demo_diagnosis_df.csv"), row.names = TRUE)
```
### MS med summaries by group
```{r demographics_meds}

@@

#####Promis x anxiety dose, group by mental vs physical health
```{r plot_Promis_x_anxiety_mental_vs_phys}
#set mood versus physical domains
mood_domains <- c("Mental_Health_and_Mood",
                "Social_Activities_Satisfaction",
                "Carrying_Out_Social_Activities",
                "Emotional_Problems", 
                "Quality_of_Life",
                "PostOp.PROMIS.Mental.Score")

@@

#save eps
postscript(file = paste0(homedir, "/results/Figures_for_msanxiety_paper/promis_with_post_op_mental_vs_physical.eps"), width = 2, height = 1.5)
  promis_with_post_op_mental_vs_physical
dev.off()
```
####Test promis*anxiety dose  stats
#####NEW JAMA ANALYSIS: Testing for Normality - all normally distributed
```{r normality_testing}
###

@@

### Lesion volumes 
#### Diagnosis - sick > healthy in vol of lesions
```{r lesion_volumes_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup() #n=123

@@

#gam lesion volumes; DiagnosisMS+Anxiety 8.262e+03  2.691e+03   3.070  0.00266 **
volume_of_mimosa_lesions_gam <- gam(volume_of_mimosa_lesions ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T) + total_volume, data = df_empi_dx_network_proportions)
summary(volume_of_mimosa_lesions_gam)

 #print summary stats; (T = 3.07, P = 0.003, Cohen's f2 =  0.354, Effect size = large)
 gam_summary_stats(volume_of_mimosa_lesions_gam)
```

@@

 #bilat anxiety dose + sex + age:  anxiety_dose  3.955e+03  1.228e+03   3.221  0.00139 **

  anxiety_dose_volume_of_mimosa_lesions_by_sex_gam <- gam(volume_of_mimosa_lesions ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T) + total_volume, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

   #print summary stats; (T = 3.221, P = 0.001, Cohen's f2 =  0.168, Effect size = medium)
 gam_summary_stats(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

@@

### Whole brain volumes 
#### Diagnosis - sick > healthy 
```{r tbv_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, total_volume) %>%
  ungroup() #n=123

@@

 #print summary stats
 gam_summary_stats(total_brain_volume_gam)

#### Anxiety dose
  - anxiety dose correlates with disease burden (total brain volume )
```{r tbv_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != -1) #n=372

@@

#### Uncinate by anxiety diagnosis (prop)
```{r uncinate_by_dx_prop}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

 #print summary stats; (T = 2.022, P = 0.045, Cohen's f2 =  0.192, Effect size = medium)
  gam_summary_stats(disease_mean_UF_prop_gam)
```

@@

#### Fx = F gives better results than FX = T
#results: Anxiety dose related to UF_R when covarying for linear effects of age and sex

@@

    #bilat anxiety dose + sex + age + total volume: anxiety_dose  0.04989    0.02397   2.082   0.0381 * 
 anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

  #print summary stats; (T = 2.082, P = 0.038, Cohen's f2 =  0.105, Effect size = small)
  gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)

@@

#### Sensitivity
###### Fornix by Anxiety Diagnosis and Dose (vol and prop, N.S)
```{r fornixByAnxietyDose}

 ##### healthy vs sick - prop
#Fornix gam Diagnosis; DiagnosisMS+Anxiety   (T = 0.67, P = 0.504, Cohen's f2 =  0.025, Effect size = small)
anxietyDiagnosis_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T) + total_volume, data = df_empi_dx_network_proportions)
summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)

#define group for dose; 
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1)

  #gam, Fornix anxiety_dose  ; anxiety_dose  2.305e-02  2.404e-02   0.959   0.3383

  anxietyGrouping_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T) + total_volume, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_F_by_sex_gam_prop)

 #print summary stats
 gam_summary_stats(anxietyDiagnosis_mean_F_by_sex_gam_prop) #"(T = 0.67, P = 0.504, Cohen's f2 =  0.025, Effect size = small)
 gam_summary_stats(anxietyGrouping_mean_F_by_sex_gam_prop) #(T = 0.959, P = 0.338, Cohen's f2 =  -0.003, Effect size = NA
```

###### UF by depGroupVar (vol/prop, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the UF findings are specific to anxiety, so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

 #mean UF, depDiagnosisMS+noDepression  T = -1.659, P = 0.098, Cohen's f2 =  0.105
dep_mean_UF_gam_prop <- gam(mean_UF_prop ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F) + total_volume, data = df_dep)
summary(dep_mean_UF_gam_prop)

  #print summary stats: (T = -1.659, P = 0.098, Cohen's f2 =  0.105, Effect size = small)
 gam_summary_stats(dep_mean_UF_gam_prop)

@@

###### JAMA ANALYSIS: Lesion & brainvol by depGroupVar (lesion vol - Y, tbv - N, N.S)
```{r UF_by_depGroupVar_lesion_tbv}
# Specificity - want to make sure the lesion and brain vol findings are nonspecific t so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

 #mean vol mimosa lesions, depDiagnosisMS+noDepression; depDiagnosisMS+noDepression -5.608e+03  1.753e+03  -3.200 0.001522 **

dep_mean_lesion_vol_gam_prop <- gam(volume_of_mimosa_lesions ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F) + total_volume, data = df_dep)
summary(dep_mean_lesion_vol_gam_prop)

  #print summary stats; #(T = -3.2, P = 0.002, Cohen's f2 =  0.178, Effect size = medium)
 gam_summary_stats(dep_mean_lesion_vol_gam_prop)

@@

###### JAMA ANALYSIS- UF Anxiety diagnosis + lesion vol
```{r uncinate_by_dx_lesion_vol}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

```
###### JAMA ANALYSIS-  UF Anxiety severity + lesion vol
```{r uncinateByAnxietyDose_lesion_vol}

@@

#cor = 0.6282121; t = 15.531, df = 370, p-value < 2.2e-16
corr_LV_to_UF_anxiety_dose <- cor.test(df_demo_and_fascicles_no_unclass_anxiety_dose$volume_of_mimosa_lesions,df_demo_and_fascicles_no_unclass_anxiety_dose$mean_UF_vol) #
  print(corr_LV_to_UF_anxiety_dose)

```
###### JAMA ANALYSIS- UF Anxiety diagnosis + total brain vol
```{r uncinate_by_dx_tbv}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

@@

###### JAMA ANALYSIS-  UF Anxiety dose + total brain volume
```{r uncinateByAnxietyDose_tbv}

@@

##### JAMA ANALYSIS- correction for multiple comparisons
```{r sensi_correction}
#uf dep group
#fornix by anxiety diag
#fornix by anxiety dose
#lesion vol anxiety diag
#lesion vol anxiety dose

anxiety_dx_vol_mimosa_lesions = summary(volume_of_mimosa_lesions_gam)$p.table[2,4]
anxiety_dose_vol_mimosa_lesions = summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)$p.table[2,4]
dep_mean_lesion_vol_gam = summary(dep_mean_lesion_vol_gam_prop)$p.table[2,4]
dep_uf_gam = summary(dep_mean_UF_gam_prop)$p.table[2,4]
anxiety_dx_fornix = summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)$p.table[2,4]
anxiety_dose_fornix = summary(anxietyGrouping_mean_F_by_sex_gam_prop)$p.table[2,4]

p_vals <- array(c(anxiety_dx_vol_mimosa_lesions,
                 anxiety_dose_vol_mimosa_lesions,
                 dep_mean_lesion_vol_gam,
                 dep_uf_gam,
                 anxiety_dx_fornix,
                 anxiety_dose_fornix))

#0.005324721 0.004567230 0.004567230 0.147282059 0.496139621 0.405577736
p_val_fdr <- p.adjust(method = "fdr", p_vals)
names(p_val_fdr) = c("anxiety_dx_vol_mimosa_lesions", "anxiety_dose_vol_mimosa_lesions", "dep_mean_lesion_vol_gam", "dep_uf_gam", "anxiety_dx_fornix", "anxiety_dose_fornix")
print(p_val_fdr)

@@

postscript(file = paste0(homedir, "/results/Figures_for_msanxiety_paper/UF_anxiety_severity_box_plot_no_jitter.eps"), width = 2, height = 1.5)
  prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot
dev.off()

@@
